SNP Analysis

### 2020-04-14 sWGS Normals.Rmd
Bug fixes for paper figures

library(tidyverse)
library(BarrettsProgressionRisk)
library(glmnet)
library(gridExtra)
library(pander)
source('lib/load_patient_metadata.R')
#source('lib/data_func.R')
#source('lib/fastPCF.R')

@@

modeldir = '~/Data/BarrettsProgressionRisk/Analysis/5e6_arms_all'

@@

rpi = read.patient.info(patient.file, demo.file)
normals = rpi$normal

@@

resids = do.call(bind_rows, purrr::map(list.files(paste0(analysis.dir, '/normals'), 'residuals.tsv', recursive = T, full.names = T), function(f) read_tsv(f)))

resids = sample.residual.variance(seg.residuals)
resids = merge(normals[,c('Patient','Hospital.Research.ID','Samplename')], resids, by.x='Samplename', by.y='sample')

failedQC = arrange(subset(resids, round(var1sd_median,3) > 0.012, select=c(Hospital.Research.ID,Samplename, var1sd_median,var1sd_sd, var1sd_Q1,var1sd_Q3)),  var1sd_median)

pander(normals[,c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')], caption='Normal samples', justify='left')
```

@@

```{r, echo=F, warning=F, fig.height=30, fig.width=8}
do.call(grid.arrange, c(plist, ncol=1))
```

@@

```{r, echo=F}

file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(dysplasia.df,coefs,plots,labels)

select.alpha = '0.9'
fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

tiled.segs = segment.matrix(tiled.normals)
tiled.arms = segment.matrix(tiled.normal.arms)

cx.score = score.cx(tiled.segs,1)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

>>>>

library(tidyverse)
library(BarrettsProgressionRisk)
library(glmnet)
library(gridExtra)
library(kableExtra)
source('lib/load_patient_metadata.R')
#source('lib/data_func.R')
#source('lib/fastPCF.R')

@@

modeldir = paste0(analysis.dir,'/models_5e6_all/50kb')

@@

rpi = read.patient.info(patient.file, demo.file)
normals = rpi$normal %>% mutate(Samplename = sub('_','-',Samplename))

@@

resids = do.call(bind_rows, purrr::map(list.files(paste0(analysis.dir, '/normals'), 'residuals.tsv', recursive = T, full.names = T), function(f) read_tsv(f)))
resids = left_join( dplyr::select(normals, Patient, Hospital.Research.ID, Samplename, Plate.Index), resids, by=c('Samplename'='samplename'))

failedQC = filter(resids, !Pass)

normals %>% dplyr::select(-matches('Cellularity|p53|Reads|Batch|Set')) %>% kable(caption = 'Normal samples') %>% kable_styling(full_width = F)
```

@@

```{r, echo=F, warning=F, fig.height=32, fig.width=12}
plist = lapply(data.files, function(f) {
  load(f)
  BarrettsProgressionRisk::plotSegmentData(segmented)
})
do.call(grid.arrange, c(plist, ncol=1))
```

@@

```{r, echo=F}
file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
fit = models[['0.9']]
lambda = performance.at.1se[['0.9']]$lambda  
cvRR = BarrettsProgressionRisk:::cvRR(dysplasia.df, coefs[['0.9']])
rm(dysplasia.df,coefs,plots,labels)
file = paste(modeldir, 'loo_0.9.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(plots, performance.at.1se, coefs, fits, pg.samp)

be.model = BarrettsProgressionRisk:::be.model.fit(fit, lambda, 50, 5e6, z.mean, z.arms.mean, z.sd, z.arms.sd, mn.cx, sd.cx, nzcoefs, cvRR)

preds = tibble()
plist = list()
for (f in data.files) { 
  load(f)

  tiles = BarrettsProgressionRisk:::tileSamples(segmented, be.model = be.model, scale=T, MARGIN=2, verbose=F)

  prr = BarrettsProgressionRisk:::predictRisk(segmented, tiles, be.model, verbose=F)

  plist[f] = grid.arrange(BarrettsProgressionRisk::copyNumberMountainPlot(prr))

  preds = bind_rows(preds, predictions(prr))
}

paste0(analysis.dir, '/normals')
files = list.files(paste0(analysis.dir, '/normals'), '5e06_.*tiled', recursive=T, full.names = T)
tiled.segs = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files(paste0(analysis.dir, '/normals'), 'arms_.*tiled', recursive=T, full.names = T)
tiled.arms = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

tiles.segs[,-1]

cx.score = BarrettsProgressionRisk::scoreCX(tiled.segs,1)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = BarrettsProgressionRisk:::unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = BarrettsProgressionRisk:::unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

### 2020-04-14 snp/SNP_Prediction_Analysis.Rmd
Bug fixes for paper figures

rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies')
pROC::ci.auc(rocBE$auc)
pROC::coords(rocBE, 'best', transpose=F)
ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)')
pROC::ci.auc(rocEX)
ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples')
ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)

# Exclude all of the above
rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP')
ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 1 only
roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1')
ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)

# Timepoint 2 only
roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2')
ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)

>>>>

rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds')
roc.plot(rocBE) + labs(title='SNP ROC, BE biopsies')

rocBE$model = 'per Biopsy'

ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE, 'SNP ROC: BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

@@

rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
roc.plot(rocEX) + labs(title='BE biopsies: Path status (HGD/IMC)')
ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocEX$model = 'HGD/IMC progressed'

@@

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
roc.plot(rocpp, 'BE biopsies: excl. low SCA samples')
ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp, 'BE biopsies: excl. low SCA samples') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)

rocpp$model = 'All excl. low SCA'

# Exclude all of the above
rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
roc.plot(rocexall, 'BE biopsies: excl. low SCA & WGD NP')
ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall, 'BE biopsies: excl. low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocexall$model = 'excl. low SCA & (NP) WGD'

# Timepoint 1 only
roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
roc.plot(roct1, 'BE biopsies: Timepoint T1')
ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1, 'BE biopsies: Timepoint T1') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)

roct1$model = 'Timepoint 1 (with exclusions)'

# Timepoint 2 only
roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds')
roc.plot(roct2, 'BE biopsies: Timepoint T2')
ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2, 'BE biopsies: Timepoint T2') + theme(text=element_text(size=12)), width=5, height=5, units='in', dpi=300)

roct2$model = 'Timepoint 2  (with exclusions)'

m = bind_rows(
      get.roc.stat(rocBE),
      get.roc.stat(rocEX),
      get.roc.stat(rocexall),
      get.roc.stat(rocpp),
      get.roc.stat(roct1),
      get.roc.stat(roct2)
) %>% dplyr::mutate_if(is.numeric, list(~round(.,2))) %>%
  dplyr::mutate(model = factor(model, levels = c(rocBE$model, rocEX$model, rocexall$model, rocpp$model, roct1$model, roct2$model), ordered = T))

dodge = position_dodge(width=0.9)
p = ggplot(m, aes(model, value, ymin=CI.min, ymax=CI.max, group=Measure)) + 
  geom_bar(aes(fill=Measure), stat='identity',position=dodge) + 
  geom_errorbar(position=dodge, width=0.3, color='black' ) +
  geom_text(aes(label=paste(value,sep=''), y=CI.max), vjust=-0.5, position=dodge) + plot.theme + 
  scale_y_continuous(limits=c(0,1)) +
  scale_fill_brewer(palette='Set2') +
  theme(legend.position = 'right',text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='ROC', title='Model comparisons') 
p 
ggsave(filename='plots/snp/roc_comparisons.png', plot=p, width=7, height=5, units='in', dpi=300)
